---
title: "Psych Replication Analysis"
author: "CJ Robinson"
date: "2025-04-23"
output: html_document
---

# Visualization and analysis for replication crisis story

See github for additional files related to scraping and the formation of the df 
referenced here

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

```{r}
library(tidyverse)
library(clipr)
library(haven)
library(ggbeeswarm)
```

```{r}
df <- read_csv("../data/replication_citation_df.csv") 

# Load the package
library(FReD)

# Load the replication dataset
fred_data <- load_fred_data()
```

# Create scatter plot

```{r}
filtered <- fred_data %>% 
  # filter out invalidated results
  filter(!is.na(result) | id == "Arb212", validated == 1) %>% 
  # highlight power posing just for reference
  mutate(result_color = ifelse(id == "Arb212", "power poser", result)) %>% 
  select(es_original, es_replication, result_color) 

rownames(filtered) <- NULL

#write out to rawgraphs
filtered %>% 
  write_clip()
```

# Make dot plot bar chart for export to AI

```{r}
df %>% 
  group_by(result) %>%
  mutate(
    idx = row_number() - 1,
    x = idx %% 5,
    y = -(idx %/% 5)
  ) %>%
  ungroup() %>% 
  ggplot(aes( x= x, y = -y, color = result)) + 
   geom_point(size = .01, ) + 
    facet_wrap(~result, scales = "free_x") 

ggsave("../img/dot_plot_bar.pdf")
```

# Field-based bar chart

```{r}
fred_data %>% 
  #misspellings in database
  group_by(discipline = ifelse(discipline == "Social Psychology", "Social psychology", discipline), result) %>% 
  tally() %>% 
  pivot_wider(names_from = result, values_from = n) %>% 
  # take only those with non null statuses and divide by total
  mutate(success_prop = `successful replication` / sum(`successful replication`, `OS not significant`, `failed replication`, na.rm = TRUE)) %>%
  # filtering over 10 successful studies for relevance
  filter(`successful replication` > 10, 
         !is.na(discipline)) %>% 
  ggplot(aes( x= reorder(discipline, desc(success_prop)), y = success_prop)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = success_prop)) + 
  coord_flip() + 
  theme_minimal()

ggsave("../img/field_bar.pdf")
```


### Data Cleaning

Do some manual cleaning for data downloaded from webpage and dates

```{r}
df_years <- df %>% 
  filter(!is.na(ref_original), 
         ref_original != "nan") %>% 
  # hand code some errors
  mutate(cited_by_count = case_when(
    ref_original == "Hart, Y., Dillon, M. R., Marantan, A., Cardenas, A. L., Spelke, E., & Mahadevan, L. (2018). The statistical shape of geometric reasoning. Scientific reports, 8(1), 1-12." ~ "0",
    ref_original == "Heck, P. R., Chabris, C. F., Watts, D. J., & Meyer, M. N. (2020). Objecting to experiments even while approving of the policies or treatments they compare. Proceedings of the National Academy of Sciences, 117(32), 18948-18950." ~ "23",
    ref_original == "Otto, R. A., Skatova, A., Madlon-Kay, S., & Daw, N. (2014). Cognitive Control Mode Predicts Behavioral Expression of Model-Based Reinforcement-Learning. In Proceedings of the Annual Meeting of the Cognitive Science Society (Vol. 36, No. 36)." ~ "0",
    .default = cited_by_count
  )) %>% 
  mutate(cited_by_count = as.numeric(cited_by_count)) %>% 
  mutate(year = str_extract(ref_original, "\\b(19|20)\\d{2}\\b")) %>% 
  group_by(ref_original) %>% 
  mutate(findings_count = n()) %>%
  ungroup() 

```

# Calculate citations

```{r}
df_years %>% 
  mutate(success_flag = ifelse(result == "Success",1,0),
         findings_flag = ifelse(findings_count == 1, "single finding", "multiple")) %>% 
  group_by(ref_original, cited_by_count,findings_flag) %>% 
  summarize(all_success = min(success_flag)) %>% 
  group_by(all_success) %>% 
  summarize(citations = sum(cited_by_count),
            paper_count = n(),
            med_citations = median(cited_by_count)) %>% 
  mutate(citations_per_paper = citations/paper_count) 
```






-----------------------------

# Other/scratch

## Tree Map

```{r}
df_years %>% 
  group_by(result) %>% 
  tally() %>%
    mutate(success_flag = ifelse(result == "Success", "Success", "Not replicable")) %>% 
  write_clip()
```



```{r}
df_years %>% 
  mutate(success_flag = ifelse(result == "Success",1,0),
         findings_flag = ifelse(findings_count == 1, "single finding", "multiple")) %>% 
  group_by(ref_original, cited_by_count,findings_flag) %>% 
  summarize(all_success = sum(success_flag),
            hyp_count = n()) %>% 
  group_by(all_success = ifelse(all_success / hyp_count > .75, 1, 0)) %>% 
  summarize(citations = sum(cited_by_count),
            paper_count = n(),
            med_citations = median(cited_by_count)) %>% 
  mutate(citations_per_paper = citations/paper_count) 
```

```{r}
df_years %>% 
  mutate(success_flag = ifelse(result == "Success",1,0),
         findings_flag = ifelse(findings_count == 1, "single finding", "multiple")) %>% 
  group_by(ref_original, cited_by_count,findings_flag) %>% 
  summarize(all_success = min(success_flag)) %>% 
  group_by(all_success, findings_flag) %>% 
  summarize(citations = sum(cited_by_count),
            paper_count = n()) %>% 
  ggplot(aes(x = factor(all_success), y= citations)) + 
  geom_bar(stat="identity") + 
  theme_minimal()
```


```{r}
df_years %>%
  group_by(year, result) %>% 
  tally() %>% 
  ggplot(aes(x =year, y = n, fill=result)) + 
  geom_bar(stat = "identity")

```

```{r}
df_years %>% 
  group_by(year, result) %>% 
  tally() %>% 
  ggplot(aes(x =year, y = n, fill=result)) + 
  geom_bar(stat = "identity", position = "fill")

```



```{r}
df_years %>% 
  group_by(year, result) %>% 
  tally() %>% 
  pivot_wider(names_from = "result", values_from = "n", names_prefix = "result.")%>% 
  mutate(across(starts_with("result."), ~replace_na(., 0))) %>% 
  mutate(sum = rowSums(across(starts_with("result.")))) %>% 
  filter(year > 2000) %>% 
  mutate(success_prop = `result.Success` / sum) %>% 
  ggplot(aes(x =year, y = 1-success_prop )) + 
  geom_bar(stat = "identity") + 
  ggtitle("Percentage of failed replications by year of publication")

```


## Citations by status 

```{r}
df_years %>% arrange(desc(cited_by_count))
```


```{r}
df_years %>% 
  mutate(success_flag = ifelse(result == "Success",1,0)) %>% 
  group_by(ref_original, cited_by_count, success_flag, result) %>% 
  summarize() %>%
  group_by(result) %>%
  summarize(citations = sum(cited_by_count),
            paper_count = n())  %>% 
  ggplot(aes(x = result, y=citations )) + 
  geom_bar(stat="identity") + 
  coord_flip() + 
  theme_minimal()
```

```{r}
df_years %>% 
  mutate(success_flag = ifelse(result == "Success",1,0)) %>% 

  group_by(ref_original, cited_by_count, success_flag, result) %>% 
  group_by(success_flag,result,findings_flag = ifelse(findings_count == 1, "single finding", "multiple")) %>%
  summarize(citations = sum(cited_by_count),
            paper_count = n())  %>% 
  ggplot(aes(x = success_flag, y= citations/paper_count)) + 
  geom_bar(stat="identity") + 
  coord_flip() + 
  facet_wrap(~findings_flag) + 
  theme_minimal()
```

```{r}
df_years %>% 
  mutate(success_flag = ifelse(result == "Success",1,0)) %>% 
  group_by(ref_original) %>% 
  mutate(findings_count = n()) %>%
  ungroup() %>% 
  group_by(ref_original, cited_by_count, success_flag, result) %>% 
  group_by(success_flag,result,findings_flag = ifelse(findings_count == 1, "single finding", "multiple")) %>%
  summarize(citations = sum(cited_by_count),
            paper_count = n())  %>% 
  ggplot(aes(x = success_flag, y= citations/paper_count)) + 
  geom_bar(stat="identity") + 
  coord_flip() + 
  facet_wrap(~findings_flag) + 
  theme_minimal()
```




```{r}
flag_ind_data <- ind_data %>% 
  mutate(
    flag = case_when(
      rowSums(!is.na(select(., starts_with("zav1.")))) > 0 ~ "cool",
      rowSums(!is.na(select(., starts_with("zav2.")))) > 0 ~ "warm",
      TRUE ~ NA_character_
    )
  )
```

```{r}
flag_ind_data %>% 
  ggplot(aes(x = as.factor(zav.dv.2), y = flag, fill = flag)) +
  geom_count(aes(size = after_stat(n)), show.legend = FALSE) +
  geom_jitter(height = 0.2, alpha = 0.5) +
  scale_size_area(max_size = 10) +
  theme_minimal() +
  labs(x = "zav.dv.2 (discrete values)", y = "Flag") +
  theme(legend.position = "top")
```

```{r}
fred_data %>% 
  group_by(discipline, result) %>% 
  tally() %>% 
  pivot_wider(names_from = result, values_from = n) %>% 
  mutate(success_prop = `successful replication` / (`successful replication` + `OS not significant` + `failed replication`)) %>%
  filter(`successful replication` > 5, 
         !is.na(discipline)) %>% 
  ggplot(aes( x= reorder(discipline, desc(success_prop)), y = success_prop)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  theme_minimal()
```

```{r}
fred_data %>% 
  filter(!is.na(result)) %>% 
  #filter(id == "Manylabs2_26") %>%
  filter(id == "Arb212") %>% 

  ggplot(aes( x= es_original, y = es_replication)) + 
  geom_point(size = .5) + 
  scale_y_continuous(limits = c(-.75, 1)) +
  scale_x_continuous(limits = c(0, 1)) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  theme_minimal()
```


```{r}
fred_data %>% 
  filter(!is.na(result) | id == "Arb212", vaidated >= 1) %>% 
  write_df
  ggplot(aes( x= es_original, y = es_replication)) + 
  geom_point(size = .5) + 
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  theme_minimal()
```


```{r}


  ggplot(aes( x= es_original, y = es_replication, color = ifelse(id == "Arb212", "power poser", result))) + 
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  geom_point(size = .5) + 
  theme_minimal()

ggsave("../img/fred_scatter.pdf")
```

```{r}
fred_data %>% 
  filter(!is.na(result) | id == "Arb212", validated >= 1) %>% 
  group_by(result) %>%
  mutate(
    idx = row_number() - 1,
    x = idx %% 5,
    y = -(idx %/% 5)
  ) %>%
  ungroup() %>% 
  ggplot(aes( x= x, y = -y, color = ifelse(id == "Arb212", "power poser", result))) + 
   geom_point(size = .01, ) + 
    facet_wrap(~result, scales = "free_x") 

ggsave("../img/dot_plot_bar.pdf")
```


